Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-21 paper.tex
fix typos

This section presents our second contribution, a framework for proving
correctness of cast representations, especially space efficient
ones. Section~\ref{sec:hypercoercion-correctness} will apply this
framework to prove the correctness of the \lazyD{} hypercoercions.

We start by defining a CEK machine \effCEK{C} that is parameterized
over the cast representation $C$. This machine is space efficient
provided that the cast representation performs compression.  The
interface between the machine and the cast representation is defined
by an abstract data type, named Cast ADT, that we define shortly.

We prove that, for any instance $C$ of the Cast ADT, \effCEK{C} is
equivelent to \ineffCEKD{] provided that $C$ satisfies a collection of
properties that we call the Lazy D Cast abstract data type.  In
Section~\ref{sec:hypercoercion-correctness} we show that the \lazyD{}
hypercoercions $H$ are an instance of the Lazy D ADT, and therefore
\effCEK{H} is equivalent to \ineffCEKD{}.

We are working on a similar theorem for \lazyUD{}.

\todo[inline]{UNDER CONSTRUCTION -JS}

@@

Fig.~\ref{fig:bisim-SC-D} defines the bisimulation relation. Here we omit rules 
that make the relation a congruence.
%
The relation $\kappa \approx \kappa$ relates two continuations. 
%
If $k \approx \kappa$, adding an identity cast on top of $k$ gives a related 
continuation. 
%
If $\kappa_1 \approx \kappa_2$, extending $\kappa_1$ and $\kappa_2$ with the 
same cast gives related results.
%
%
The relation $v \approx v : T$ relates two value and also report their type. We 
will see shortly why reporting the type. 
%
If two values of type $P$ are related, injecting them to $\TOOdyn$ gives 
related values.
%
Unit values are of course related.
%
To relate to a non-casted function in \ineffCEKD, the casts in the \effCEK{C} 
function must be identity.
%
If functions are related, we can apply the same cast on both side.
%
Rules for products and sums are similar.

\todo[inline]{In Fig.~\ref{fig:bisim-SC-D}, how to suppress rule numbers like 
(1), (2), etc.? -KC}
\begin{figure}
%	\fbox{$\text{\effCEK{C}} \approx \text{\ineffCEKD}$}
	\[\dots\]

	\fbox{$\kappa \approx \kappa$}
	\begin{gather*}
	\inference{
		k \approx \kappa
	}{
		wrap(k) \approx \kappa
	}
	\quad
	\inference{
		\kappa_1 \approx \kappa_2
	}{
		ext(cast(T_1,l,T_2),\kappa_1) \approx 
		\kOOcast{\cOOcast{T_1}{l}{T_2}}{\kappa_2}
	}
\\
	\end{gather*}
\\	
	\fbox{$v \approx v : T$}
	\begin{gather}
	\inference{
		v_1 \approx v_2 : P
	}{
		\hcvOOinj{P}{v_1} \approx \vOOcast{v_2}{\cOOcast{P}{l}{\TOOdyn}}
		: \TOOdyn
	}
\\ \\
	\inference{}{
		\hcvOOtt \approx \vOOtt : \POOunit
	}
	\quad
	\inference{}{
		\hcvOOtt \approx \vOOcast{\vOOtt}{\cOOcast{\POOunit}{l}{\POOunit}}
		: \POOunit
	}
	\\ \\
	\inference{
		\rho_1 \approx \rho_2 &
		dom(\rho_1) = dom(\rho_2) = \Gamma \\
		\Gamma , x : T_1 \vdash e : T_2
	}{
		\hcvOOfun{id(T_1)}{\rho_1}{x}{e}{id(T_2)}
		\approx
		\vOOfun{\rho_2}{x}{e}
		: \POOfun{T_1}{T_2}
	}
\\ \\
\inference{
\hcvOOfun{c_1}{\rho}{x}{e}{c_2} \approx v_2 : \POOfun{T_1}{T_2} \\
%		\judgeType{\hcvOOfun{c_1}{\rho}{x}{e}{c_2}}{\POOfun{T_1}{T_2}} \\
c_1' = seq(cast(T_3,l,T_1),c_1) \\
c_2' = seq(c_2,cast(T_2,l,T_4))
}{
\hcvOOfun{c_1'}{\rho}{x}{e}{c_2'}
\approx
\vOOcast{v}{\cOOcast{\POOfun{T_1}{T_2}}{l}{\POOfun{T_3}{T_4}}}
: \POOfun{T_3}{T_4}
}
\\ \\
	\inference{
		v_1 \approx v_2 : T_1 &
		v_3 \approx v_4 : T_2 &
	}{
		\hcvOOcons{v_1}{id(T_1)}{v_3}{id(T_2)}
		\approx
		\vOOcons{v_2}{v_3}
		: \POOprod{T_1}{T_2}
	}
\\ \\
	\inference{
		\hcvOOcons{v_1}{c_1}{v_3}{c_2} \approx v_2 : \POOsum{T_1}{T_2} \\
		c_1' = seq(c_1,cast(T_1,l,T_3)) \\
		c_2' = seq(c_2,cast(T_2,l,T_4))
	}{
		\hcvOOcons{v_1}{c_1'}{v_3}{c_2'} \approx
		\vOOcast{v}{\cOOcast{\POOprod{T_1}{T_2}}{l}{\POOprod{T_3}{T_4}}}
		: \POOprod{T_3}{T_4}
	}
\\ \\
	\inference{
		v_1 \approx v_2 : T
	}{
		\hcvOOinl{v_1}{id(T)} \approx \vOOinl{v_2}
		: \POOsum{T}{T'}
	}
	\quad
	\inference{
		\hcvOOinl{v_1}{c} \approx v_2
		: \POOsum{T_1}{T_2} &
		c' = seq(c,cast(T_1,l,T_3))
	}{
		\hcvOOinl{v_1}{c'} \approx
		\vOOcast{v}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
		: \POOsum{T}{T'}	
	}
	\\
	\inference{
	v_1 \approx v_2 : T
	}{
	\hcvOOinr{v_1}{id(T)} \approx \vOOinr{v_2}
	: \POOsum{T'}{T}
	}
	\quad
	\inference{
	\hcvOOinr{v_1}{c} \approx v_2
	: \POOsum{T_1}{T_2} &
	c' = seq(c,cast(T_2,l,T_4))
	}{
	\hcvOOinr{v_1}{c'} \approx
	\vOOcast{v}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
	: \POOsum{T'}{T}
	}
	\end{gather}
	\caption{Bisimulation between \effCEK{C} and \ineffCEKD}
	\label{fig:bisim-SC-D}
\end{figure}

>>>>

This section presents our second contribution, a framework for proving
correctness of cast representations. The next section will apply this
framework to prove the correctness of the \lazyD{} hypercoercions.

We start by defining a CEK machine \effCEK{C} that is parameterized
over the cast representation $C$. This machine is space efficient
provided that the cast representation performs compression.  The
interface between the machine and the cast representation is defined
by an abstract data type, named Cast ADT, that we define shortly.
%

\todo[inline]{UNDER CONSTRUCTION. -JS}

If $C$, an implementation of the Cast ADT inteface, satisifies a collection of 
properties that we defined as Lazy D Cast ADT. We can instantiate our framework 
theorem that we define shortly to prove that evaluating an program in 
\effCEK{C} gives the same result as \ineffCEKD. We are working on a
similar theorem for \lazyUD{}.

\todo[inline]{UNDER CONSTRUCTION. -KC}

The framework is a theorem saying that for all cast representation $C
$, if $C$ satisfies a collection of properties, then it would
respect the Lazy D semantics of cast calculi. We are working on a
similar theorem for \lazyUD{}.

%then instantiating our parameterized space-efficient interpreter with$C$
%gives an interpter that is extensionally equal to the standard interpreter for 
%\lazyD\ cast calculus.

@@

\todo[inline]{In Fig.~\ref{fig:bisim-SC-D}, how to suppress rule numbers like 
	(1), (2), etc.? -KC}
Fig.~\ref{fig:bisim-SC-D} defines the bisimulation relation. Here we omit rules 
that make the relation a congruence.
%
The relation $\kappa \approx \kappa$ relates two continuations. 
%
If $k \approx \kappa$, adding an identity cast on top of $k$ gives a related 
continuation. 
%
If $\kappa_1 \approx \kappa_2$, extending $\kappa_1$ and $\kappa_2$ with the 
same cast gives related results.
%
%
The relation $v \approx v : T$ relates two value and also report their type. We 
will see shortly why reporting the type. 
%
If two values of type $P$ are related, injecting them to $\TOOdyn$ gives 
related values.
%
Unit values are of course related.
%
To relate to a non-casted function in \ineffCEKD, the function in \effCEK{C} 
must use identities as its casts.
%
If functions are related, we can apply the same cast on both side.
%
Rules for products and sums are similar.

\begin{lemma}[SoAndSo]
	Assume $C$ implements Lazy D Cast ADT
	and $v_1$ is a value in \effCEK{C}
	and $v_2$ is a value in \ineffCEKD\
	and $v_1 \approx v_2 : T_1$,
	\[
	applyCast_{\text{\effCEK{C}}}(v,cast(T_1,l,T_2)) 
	\approx 
	applyCast_{\text{\ineffCEKD}}(v,\cOOcast{T_1}{l}{T_2})
	\]
\end{lemma}
\begin{proof}
	Case splitting whether $T_1$ is shallowly consistent with $T_2$.
	If yes, case splitting how they are shallowly consistent.
	Then apply properties (3) - (11) of Lazy D Cast when applicable.
\end{proof}

\begin{figure}
	$\text{\effCEK{C}} \approx \text{\ineffCEKD}$
	\[\dots\]

	\fbox{$\kappa \approx \kappa$}
	\begin{gather*}
	\inference{
		k \approx \kappa
	}{
		wrap(k) \approx \kappa
	}
	\quad
	\inference{
		\kappa_1 \approx \kappa_2
	}{
		ext(cast(T_1,l,T_2),\kappa_1) \approx 
		\kOOcast{\cOOcast{T_1}{l}{T_2}}{\kappa_2}
	}
\\
	\end{gather*}
\\	
	\fbox{$v \approx v : T$}
	\begin{gather}
	\inference{
		v_1 \approx v_2 : P
	}{
		\hcvOOinj{P}{v_1} \approx \vOOcast{v_2}{\cOOcast{P}{l}{\TOOdyn}}
		: \TOOdyn
	}
\\ \\
	\inference{}{
		\hcvOOtt \approx \vOOtt : \POOunit
	}
	\quad
	\inference{}{
		\hcvOOtt \approx \vOOcast{\vOOtt}{\cOOcast{\POOunit}{l}{\POOunit}}
		: \POOunit
	}
	\\ \\
	\inference{
		\rho_1 \approx \rho_2 &
		dom(\rho_1) = dom(\rho_2) = \Gamma \\
		\Gamma , x : T_1 \vdash e : T_2
	}{
		\hcvOOfun{id(T_1)}{\rho_1}{x}{e}{id(T_2)}
		\approx
		\vOOfun{\rho_2}{x}{e}
		: \POOfun{T_1}{T_2}
	}
\\ \\
\inference{
\hcvOOfun{c_1}{\rho}{x}{e}{c_2} \approx v_2 : \POOfun{T_1}{T_2} \\
%		\judgeType{\hcvOOfun{c_1}{\rho}{x}{e}{c_2}}{\POOfun{T_1}{T_2}} \\
c_1' = seq(cast(T_3,l,T_1),c_1) \\
c_2' = seq(c_2,cast(T_2,l,T_4))
}{
\hcvOOfun{c_1'}{\rho}{x}{e}{c_2'}
\approx
\vOOcast{v}{\cOOcast{\POOfun{T_1}{T_2}}{l}{\POOfun{T_3}{T_4}}}
: \POOfun{T_3}{T_4}
}
\\ \\
	\inference{
		v_1 \approx v_2 : T_1 &
		v_3 \approx v_4 : T_2 &
	}{
		\hcvOOcons{v_1}{id(T_1)}{v_3}{id(T_2)}
		\approx
		\vOOcons{v_2}{v_3}
		: \POOprod{T_1}{T_2}
	}
\\ \\
	\inference{
		\hcvOOcons{v_1}{c_1}{v_3}{c_2} \approx v_2 : \POOsum{T_1}{T_2} \\
		c_1' = seq(c_1,cast(T_1,l,T_3)) \\
		c_2' = seq(c_2,cast(T_2,l,T_4))
	}{
		\hcvOOcons{v_1}{c_1'}{v_3}{c_2'} \approx
		\vOOcast{v}{\cOOcast{\POOprod{T_1}{T_2}}{l}{\POOprod{T_3}{T_4}}}
		: \POOprod{T_3}{T_4}
	}
\\ \\
	\inference{
		v_1 \approx v_2 : T_1
	}{
		\hcvOOinl{v_1}{id(T_1)} \approx \vOOinl{v_2}
		: \POOsum{T_1}{T_2}
	}
	\quad
	\inference{
		\hcvOOinl{v_1}{c} \approx v_2
		: \POOsum{T_1}{T_2} &
		c' = seq(c,cast(T_1,l,T_3))
	}{
		\hcvOOinl{v_1}{c'} \approx
		\vOOcast{v}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
		: \POOsum{T_3}{T_4}	
	}
	\\
	\inference{
	v_1 \approx v_2 : T_2
	}{
	\hcvOOinr{v_1}{id(T_2)} \approx \vOOinr{v_2}
	: \POOsum{T_1}{T_2}
	}
	\quad
	\inference{
	\hcvOOinr{v_1}{c} \approx v_2
	: \POOsum{T_1}{T_2} &
	c' = seq(c,cast(T_2,l,T_4))
	}{
	\hcvOOinr{v_1}{c'} \approx
	\vOOcast{v}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
	: \POOsum{T_3}{T_4}
	}
	\end{gather}
	\caption{Bisimulation between \effCEK{C} and \ineffCEKD}
	\label{fig:bisim-SC-D}
\end{figure}
